Regional Inequalities and economic growth; cross section and times series analysis

### 2023-12-07 MSB104_Submission_Group1.qmd

**Expected Inverted U- and N shaped curves**

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

@@

@@

@@

```{r}
#| echo: false
#| output: false
#| fig-cap: Regression statistics of all countries seperately for the year 2010

@@

@@

```{r}
#| echo: false
#| output: false
#| label: fig_scatterplot-all
#| cap-location: bottom
#| fig-cap: "Relationship Between GDP per Capita and Gini"

@@

@@

```{r}
#| echo: false
#| output: false
#| fig-cap: Multiple Linear Regression Model

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

```{r}
#| echo: false

@@

@@

```{r}
#| echo: false

@@

# Panel estiamtion

plm1 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("id_nuts0"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

plm2 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("year"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

plm3 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("id_nuts2"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

plm4 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = gini_filtered, index = c("year","id_nuts2"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

@@

@@

>>>>

**Expected Inverted U- and N shaped curves (er dette riktig?!?, se tilbakemelding)**

@@

```{r}
#| echo: false
#| label: fig-1
#| fig-cap: "GDP per capita Portugal"

@@

```{r}
#| echo: false
#| #| label: fig-2
#| fig-cap: "GDP per capita France"

@@

```{r}
#| echo: false
#| #| label: fig-3
#| fig-cap: "GDP per capita Hungary"

@@

```{r}
#| echo: false
#| label: fig-4
#| fig-cap: "GDP per capita Slovakia"

@@

```{r}
#| echo: false
#| label: fig-5
#| fig-cap: "GDP per capita Denmark"

@@

```{r}
#| echo: false
#| label: fig-6
#| fig-cap: "Bar chart"

@@

```{r}
#| echo: false
#| label: fig-7
#| fig-cap: "Time-series plot"

@@

Ska me ha med denne? er det egentlig nødvendig??

```{r}
#| echo: false
#| output: false
#| label: fig-8
#| fig-cap: "Box plot"

@@

bør kanskje skrive noe mer om hva gini er, tror det står på side 116 eller 106 i artikkel

@@

```{r}
#| echo: false
#| output: false
#| label: tbl-1
#| fig-cap: "Regression analysis all countries combined for the year 2010"

@@

```{r}
#| echo: false
#| output: false
#| label: tbl-2
#| fig-cap: "Regression statistics of all countries seperately for the year 2010"

@@

```{r}
#| echo: false
#| output: false
#| label: fig-9
#| fig-cap: "Residuals vs fitted plot and normal Q-Q residuals plot"

@@

```{r}
#| echo: false
#| output: false
#| label: fig-10
#| cap-location: bottom
#| fig-cap: "Relationship Between GDP per Capita and Gini"

@@

MÅ SKRIVE MER HER!!

@@

```{r}
#| echo: false
#| output: false
#| label: tbl-3
#| fig-cap: "Multiple Linear Regression Model"

@@

```{r}
#| echo: false
#| #| label: tbl-4
#| fig-cap: "Regression analysis France vs other countries"

@@

```{r}
#| echo: false
#| label: tbl-5
#| fig-cap: "Linear, logarithmic and quadratic model"

@@

```{r}
#| echo: false
#| label: fig-11
#| fig-cap: "Figure of logarithmic model for France: Predicted vs actual gini coefficients"

@@

```{r}
#| echo: false
#| label: fig-12
#| fig-cap: "Figure of logarithmic model for other countries: Predicted vs actual gini coefficients"

@@

```{r}
#| echo: false
#| label: fig-13
#| fig-cap: "Figure of quuadratic model for France: Predicted vs actual gini coefficients"

@@

```{r}
#| echo: false
#| label: fig-14
#| fig-cap: "Figure of quuadratic model for the other countries: Predicted vs actual gini coefficients"

@@

MÅ FIKSE PÅ NOE MED N OG U SHAPE (SE TILBAKEMELDING)

@@

```{r}
#| echo: false
#| label: fig-15
#| fig-cap: "Residuals vs fitted values for linear model, all countries combined"

@@

```{r}
#| echo: false
#| output: false
# filtering data 
fgini_nuts_2 <- gini_nuts_2 |> 
  filter(!is.na(population),
         population != 0,
         GDP_per_capita != 0, 
         GDP_per_capita != "NA"
  ) 
```

```{r}
#| echo: false
#| output: false
#| label: tbl-6
#| fig-cap: "Panel estimations"

# Panel estimations

plm1 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("id_nuts0"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

plm2 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("year"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

plm3 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("id_nuts2"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

plm4 <- plm(gini_n2 ~ GDP_per_capita + log(GDP_per_capita), data = fgini_nuts_2, index = c("year","id_nuts2"), model = "within",effect = "twoway") |> 
  coeftest(vcov. = vcovHC, type = "HC1", save = TRUE)

@@

By looking at the results of the panel estimation, we can see a distinct relationship between GDP per capita and regional inequality, with a significant variability observed across the different models. Model 1 and 2 shows a significant positive correlation between GDP per capita and regional inequality, which suggests that higher economic output per capita is associated with greater inequality. The negatice coefficients for the log of GDP per capita imply that relative increases in economic prosperity could lead to an reduction in inequality; however, models 3 and 4 show that this association becomes less strong or irrelevant, suggesting that the influence of economic forces on inequality depends on regional or temporal characteristics.

These variations underscore the influence of fixed effects, showing that country-specific, temporal and regional characteristics can alter our understanding of economic disparity. When we compare it to cross-sectional estimates, we can see the panel data's ability to caputre temporal changes and control for heterogenity, offers a more comprenehsive perspective. This is in contrast with the static nature of cross-sectional analysis, which might overlook the complexities and nature of economic inequality. The panel approach provides deeper insights into the development of regional disparities and factors influencing them.

@@

The findings from the panel estimation offer insightful revelations about the relationship between GDP per capita and regional inequality; but we also have to look at potential methodological limitations and biases. The models sensitivity to different fixed effects suggests a nuanced, but potentially inconsistent understanding of inequality. This variability reflect the complexity of economic factors at play, and raises concerns about omitted variable bias, where key inlfuencers like industrial composition or demographic changes might be missing.

The inherent challenges in accurately measuring GDP per capita and inequality could introduce measurement errors, affecting the reliability of the results. While panel data provides a dynamic view over cross-sectional data, establishing causality remains elusive, and the correlations observed should be interpreted with caution. The study's generalizability is also limited with its focus on a specific set of countries and times, which may not represent broader global dynamics. There is a low R-squared in some of the models, something that might've occured due to unexplained variances, suggesting that other significant, but unaccounted factors might be influencing regional inequality. Additionally, the economic significance on the finding needs careful consideration in order to understand the-world impact of these economic relationships. Overall the findings should be interpreted within the context of these potential biases and limitations, and they underscore the complexity of modeling and understanding economic phenomenas such as regional inequality.
